---
title: "Data2002 Project 1"
date: "`r Sys.Date()`"
author: "541035662 530029720 540872536 550505479 530210568"
format: 
  html: 
    embed-resources: true # Creates a single HTML file as output
    code-fold: true # Code folding; allows you to show/hide code chunks
    code-tools: true # Includes a menu to download the code file 
    # code-tools are particularly important if you use inline R to 
    # improve the reproducibility of your report
table-of-contents: true # (Optional) Creates a table of contents
number-sections: true # (Optional) Puts numbers next to heading/subheadings
---

# Introduction

[Motivation + client]

[Dataset description]

[Survey questions used]

[Potential biases and limitations of the survey]


## Data Wrangling

All data processing for this report was conducted in R (R Core Team 2026) within the Rmarkdown environment (Xie, Allaire, and Grolemund 2018) to ensure reproducibility. Data wrangling was performed using base R and the tidyverse (Wickham et al. 2019). Visualisations were produced with ggplot2 (Wickham et al. 2019).

The raw survey dataset was imported and variable names were cleaned and shortened to improve readability throughout the analysis. Several categorical variables were recoded for clarity, including the assess variable, where “Other” responses were removed to retain only meaningful categories. The multitask variable was also converted into a binary factor with consistent labels.

Missing values were not removed globally; instead, they were handled on a case‑by‑case basis to preserve as much information as possible.

A threshold was applied to the work_hrs variable to remove implausible values above 70 hours per week. This cutoff was chosen because working more than 70 hours per week exceeds a standard full‑time workload and leaves insufficient time for sleep, commuting, and university study. Boxplots were generated to compare the distribution of weekly employment hours before and after cleaning, illustrating the impact of this threshold. The cleaned dataset was then saved for subsequent analysis.


```{r}
library(tidyverse)

raw <- read_csv("DATA2X02_2026_2_survey_clean.csv", show_col_types = FALSE)

df <- raw %>%
  rename(
    age          = `What is your age? (years)`,
    height       = `What is your approximate height? (cms)`,
    intl_dom     = `Are you an International or Domestic student?`,
    unit         = `Are you in DATA2002 or DATA2902?`,
    grade1001    = `What grade did you get in DATA1001/1901 (or equivalent unit)?`,
    gradehope    = `What grade do you hope to get in DATA2002/2902?`,
    assess       = `When it comes to uni assessments, what do you tend to do?`,
    commute      = `On average, how long does it take you to get to campus? (mins) [1 direction]`,
    pay_rent     = `Do you pay rent?`,
    rent_amt     = `On average, how much rent do you pay per week? (dollars)`,
    work_hrs     = `On average, how many hours a week do you spend in paid employment? (hours)`,
    sleep_hrs    = `On average, how many hours sleep do you get per night? (hours)`,
    music        = `What is your favourite type of music?`,
    aliens       = `Do you believe in the existence of aliens? (Extraterrestial life)`,
    water        = `On average, how much water do you drink in a day? (litres)`,
    n_lang       = `How many spoken languages can you speak fluently?`,
    multitask    = `Do you multi-task in lectures? (ie do other tasks concurrently)`
  ) %>%
  filter(assess != "Other") %>%
  mutate(assess = factor(assess, levels = c("Immediately", "Gradually", "Cram")),
         multitask = factor(multitask, levels = c("no", "yes"), labels = c("No", "Yes")))

threshold <- 70

df_clean <- df %>% filter(work_hrs <= threshold)

ggplot(bind_rows(
    df %>% mutate(stage = "Before cleaning"),
    df_clean %>% mutate(stage = "After cleaning")
  ),
  aes(x = stage, y = work_hrs, fill = stage)) +
  geom_boxplot() +
  scale_fill_manual(values = c("Before cleaning" = "#8ecae6", "After cleaning" = "#ffb703")) +
  labs(
    title = "Weekly employment hours before and after cleaning",
    x = NULL, y = "Weekly employment hours", fill = NULL
  ) +
  theme_minimal()

write_csv(df_clean, "survey_cleaned.csv")
```

# Results

## RQ1

[Motivation]

[EDA figure + summary statistics]

[Hypotheses]

[Assumptions]

[Test statistic]

[Observed test statistic]

[p-value]

[Decision + contextual conclusion]

[Limitations]


## RQ2
### Step 1: Data import & cleaning (multitasking vs cramming)

```{r}
#| label: rq2-setup
#| message: false
library(tidyverse)
```

```{r}
#| label: rq2-clean
survey <- read_csv("survey_cleaned.csv")

# Keep only the variables needed for RQ2, and derive a binary
# "Cram vs Not Cram" version of `assess` so the test lines up with the RQ
# statement ("multitasking and cramming"), rather than the original 3-level
# assessment-style variable (Gradually / Immediately / Cram).
rq2_data <- survey %>%
  select(assess, multitask) %>%
  drop_na(assess, multitask) %>%
  mutate(
    assess    = factor(assess, levels = c("Gradually", "Immediately", "Cram")),
    is_cram   = if_else(assess == "Cram", "Cram", "Not Cram"),
    is_cram   = factor(is_cram, levels = c("Not Cram", "Cram")),
    multitask = factor(multitask, levels = c("No", "Yes"))
  )

glimpse(rq2_data)
```

We only select `assess` and `multitask` (not the whole cleaned dataset), and
only drop rows missing on these two columns — so this subsetting is specific
to RQ2 and doesn't discard rows that are unusable for other RQs. `assess` is
collapsed from 3 levels into a binary `is_cram` variable to match the RQ2
question directly (cram vs not cram).

### Step 2: Exploratory visualisation (multitasking vs cramming)

```{r}
#| label: rq2-eda
#| fig-cap: "Proportion of students who multitask, by cramming behaviour"
rq2_data %>%
  count(is_cram, multitask) %>%
  group_by(is_cram) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = is_cram, y = prop, fill = multitask)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "Assessment preparation style",
    y = "Proportion of students",
    fill = "Multitasks?",
    title = "Multitasking is more common among students who cram"
  ) +
  theme_minimal(base_size = 13)
```

Supplementary (optional) — the same relationship broken down by the original
3-level `assess` variable, useful as a descriptive figure even though the
formal test uses the collapsed 2x2 version:

```{r}
#| label: rq2-eda-3level
survey %>%
  drop_na(assess, multitask) %>%
  count(assess, multitask) %>%
  group_by(assess) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = assess, y = prop, fill = multitask)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Assessment preparation style (3 levels)",
       y = "Proportion of students", fill = "Multitasks?") +
  theme_minimal(base_size = 13)
```

Both plots show the same pattern: the multitasking rate is visibly higher
among students who cram (~75%) than among those who don't (~52% "Gradually",
~46% "Immediately"). This motivates formally testing the association in the
next step.

### Step 3: Hypothesis test — Chi-squared test of independence

**H:**
$H_0$: multitasking behaviour is independent of cramming behaviour.
$H_1$: multitasking behaviour is *not* independent of cramming behaviour.

```{r}
#| label: rq2-table
cram_tab <- table(rq2_data$is_cram, rq2_data$multitask)
cram_tab
```

**A (assumptions):** The chi-squared test of independence requires (i)
independent observations (one row per student — reasonable here, no student
appears twice in the survey) and (ii) expected cell counts $\geq 5$ in every
cell of the contingency table.

```{r}
#| label: rq2-assumptions
chi_test <- chisq.test(cram_tab, correct = FALSE)
chi_test$expected
```

All four expected counts are well above 5 (minimum ≈ 51.5), so the
chi-squared approximation is valid here — there is no need to fall back on
Fisher's exact test or a Yates continuity correction.

**T & P (test statistic & p-value):**

```{r}
#| label: rq2-test
chi_test
```

```{r}
#| label: rq2-effect-size
# Effect size (Cramer's V), to accompany the p-value
n <- sum(cram_tab)
cramers_v <- sqrt(unname(chi_test$statistic) / n)
cramers_v
```

Observed: $\chi^2 = 21.35$, $df = 1$, $p \approx 3.8\times10^{-6}$,
Cramér's $V \approx 0.18$ (small-to-moderate association).

**C (conclusion with limitations):**
There is very strong evidence ($p < 0.001$) against the null hypothesis of
independence: students who cram are considerably more likely to multitask
(75.2%) than students who don't (51.8%). However:

1. This is an observational, self-selected survey sample — voluntary
   response bias means busier or more disorganised students may be
   over- or under-represented.
2. "Cram" and "multitask" are both self-reported and self-interpreted
   categories, which may not be applied consistently across respondents.
3. Association is not causation — a third factor (e.g. workload,
   procrastination tendency) could plausibly drive both behaviours.
4. Collapsing `assess` from 3 levels to 2 (Cram vs Not Cram) loses some
   information about the distinction between "Gradually" and
   "Immediately" preparers.


## RQ3: Do students who multitask during lectures spend a different average number of hours per week in paid employment compared with students who do not multitask?

Paid employment represents an external time commitment for students. Students balancing more paid work with university study may exhibit different lecture behaviours. Having examined paid employment in relation to cramming and multitasking in relation to cramming, we investigate whether paid-employment hours also differ between students who do and do not multitask during lectures.

```{r}
#| label: tbl-flipper
#| tbl-cap: "Summary statistics of weekly paid employment hours by multitasking status."

rq3_summary <- df_clean %>%
  filter(!is.na(work_hrs), !is.na(multitask)) %>%
  group_by(multitask) %>%
  summarise(
    n = n(),
    mean = mean(work_hrs),
    median = median(work_hrs),
    sd = sd(work_hrs),
    Q1 = quantile(work_hrs, 0.25),
    Q3 = quantile(work_hrs, 0.75),
    min = min(work_hrs),
    max = max(work_hrs)
  )

knitr::kable(rq3_summary)
```
The boxplot in @fig-work-multitask compares weekly paid-employment hours between students who reported multitasking during lectures and those who did not. Students who reported multitasking had a higher mean weekly paid-employment time (8.02 hours) than those who did not multitask (6.23 hours), giving an observed mean difference of 1.79 hours per week. Both groups show visually right-skewed distributions, with many students reporting few or no hours of paid employment and a smaller number reporting substantially higher working hours. This observed difference in group means is examined formally using a permutation test.
```{r}
#| label: fig-work-multitask
#| fig-cap: "Distribution of weekly paid employment hours by multitasking status."

df_clean %>%
  filter(!is.na(work_hrs), !is.na(multitask)) %>%
  ggplot(aes(x = multitask, y = work_hrs)) +
  
  geom_boxplot(
    aes(fill = multitask),
    width = 0.5,
    outlier.shape = NA,
    alpha = 0.4
  ) +
  
  geom_jitter(
    aes(colour = "Individual observations"),
    position = position_jitter(
      width = 0.15,
      seed = 123
    ),
    alpha = 0.35,
    size = 1.5
  ) +
  
  stat_summary(
    aes(colour = "Group mean"),
    fun = mean,
    geom = "point",
    shape = 18,
    size = 4
  ) +
  
  stat_summary(
    fun = mean,
    geom = "text",
    aes(label = after_stat(round(y, 2))),
    vjust = -0.75,
    size = 3.5
  ) +
  
  scale_fill_manual(
    name = "Multitasking",
    values = c(
      "No" = "#A8DADC",
      "Yes" = "#F4A261"
    )
  ) +
  
  scale_colour_manual(
    name = NULL,
    values = c(
      "Individual observations" = "#457B9D",
      "Group mean" = "red"
    )
  ) +
  
  labs(
    x = "Multitasking during lectures",
    y = "Weekly paid employment (hours)",
    title = "Weekly paid employment hours by multitasking status",
    caption = "Diamonds and labels indicate group means."
  ) +
  
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "top"
  )
```

[Hypotheses]

[Permutation procedure]

[Permutation null distribution figure]

[Permutation p-value]

[Decision + contextual conclusion]

[Limitations]


# Conclusion

[Integrate the three findings into one story]

[Implications for client]

[Overall limitations / no causal inference]


# AI Usage Statement

[...]


# ISI Professional Ethics Statement

[...]


# References