---
title: "Data2002 Project 1"
date: "`r Sys.Date()`"
author: "541035662 530029720 540872536 550505479 530210568"
format: 
  html: 
    embed-resources: true # Creates a single HTML file as output
    code-fold: true # Code folding; allows you to show/hide code chunks
    code-tools: true # Includes a menu to download the code file 
    # code-tools are particularly important if you use inline R to 
    # improve the reproducibility of your report
table-of-contents: true # (Optional) Creates a table of contents
number-sections: true # (Optional) Puts numbers next to heading/subheadings
---

## Introduction

<See Assignment Briefing Page>

## Problem & Client

<See Assignment Briefing Page>

## Dataset

<See Assignment Briefing Page>

## Biases

<See Assignment Briefing Page>

## Data cleaning

**Load, rename, and clean**

Import the survey dataset, rename variables for clarity, remove "Other" responses from `assess`, and remove `work_hrs` values above 70 hours per week. This threshold was chosen because more than 70 hours is equivalent to a 10-hour day with no days off, exceeding a standard full-time work week and leaving little time for sleep, classes, or commuting.

**Compare and save**

Boxplot `work_hrs` before and after cleaning, then save the cleaned dataset as `survey_cleaned.csv`.

```{r}
library(tidyverse)

raw <- read_csv("DATA2X02_2026_2_survey_clean.csv", show_col_types = FALSE)

df <- raw %>%
  rename(
    age          = `What is your age? (years)`,
    height       = `What is your approximate height? (cms)`,
    intl_dom     = `Are you an International or Domestic student?`,
    unit         = `Are you in DATA2002 or DATA2902?`,
    grade1001    = `What grade did you get in DATA1001/1901 (or equivalent unit)?`,
    gradehope    = `What grade do you hope to get in DATA2002/2902?`,
    assess       = `When it comes to uni assessments, what do you tend to do?`,
    commute      = `On average, how long does it take you to get to campus? (mins) [1 direction]`,
    pay_rent     = `Do you pay rent?`,
    rent_amt     = `On average, how much rent do you pay per week? (dollars)`,
    work_hrs     = `On average, how many hours a week do you spend in paid employment? (hours)`,
    sleep_hrs    = `On average, how many hours sleep do you get per night? (hours)`,
    music        = `What is your favourite type of music?`,
    aliens       = `Do you believe in the existence of aliens? (Extraterrestial life)`,
    water        = `On average, how much water do you drink in a day? (litres)`,
    n_lang       = `How many spoken languages can you speak fluently?`,
    multitask    = `Do you multi-task in lectures? (ie do other tasks concurrently)`
  ) %>%
  filter(assess != "Other") %>%
  mutate(assess = factor(assess, levels = c("Immediately", "Gradually", "Cram")))

threshold <- 70

df_clean <- df %>% filter(work_hrs <= threshold)

ggplot(bind_rows(
    df %>% mutate(stage = "Before cleaning"),
    df_clean %>% mutate(stage = "After cleaning")
  ),
  aes(x = stage, y = work_hrs, fill = stage)) +
  geom_boxplot() +
  scale_fill_manual(values = c("Before cleaning" = "#8ecae6", "After cleaning" = "#ffb703")) +
  labs(
    title = "Weekly employment hours before and after cleaning",
    x = NULL, y = "Weekly employment hours", fill = NULL
  ) +
  theme_minimal()

write_csv(df_clean, "survey_cleaned.csv")
```

## RQ1

<See Assignment Briefing page>

## RQ2

<See Assignment Briefing page>
## RQ2 — Step 1: Data import & cleaning (multitasking vs cramming)

```{r}
#| label: rq2-setup
#| message: false
library(tidyverse)
```

```{r}
#| label: rq2-clean
survey <- read_csv("survey_cleaned.csv")

# Keep only the variables needed for RQ2, and derive a binary
# "Cram vs Not Cram" version of `assess` so the test lines up with the RQ
# statement ("multitasking and cramming"), rather than the original 3-level
# assessment-style variable (Gradually / Immediately / Cram).
rq2_data <- survey %>%
  select(assess, multitask) %>%
  drop_na(assess, multitask) %>%
  mutate(
    assess    = factor(assess, levels = c("Gradually", "Immediately", "Cram")),
    is_cram   = if_else(assess == "Cram", "Cram", "Not Cram"),
    is_cram   = factor(is_cram, levels = c("Not Cram", "Cram")),
    multitask = factor(multitask, levels = c("no", "yes"))
  )

glimpse(rq2_data)
```

We only select `assess` and `multitask` (not the whole cleaned dataset), and
only drop rows missing on these two columns — so this subsetting is specific
to RQ2 and doesn't discard rows that are unusable for other RQs. `assess` is
collapsed from 3 levels into a binary `is_cram` variable to match the RQ2
question directly (cram vs not cram).

## RQ2 — Step 2: Exploratory visualisation (multitasking vs cramming)

```{r}
#| label: rq2-eda
#| fig-cap: "Proportion of students who multitask, by cramming behaviour"
rq2_data %>%
  count(is_cram, multitask) %>%
  group_by(is_cram) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = is_cram, y = prop, fill = multitask)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "Assessment preparation style",
    y = "Proportion of students",
    fill = "Multitasks?",
    title = "Multitasking is more common among students who cram"
  ) +
  theme_minimal(base_size = 13)
```

Supplementary (optional) — the same relationship broken down by the original
3-level `assess` variable, useful as a descriptive figure even though the
formal test uses the collapsed 2x2 version:

```{r}
#| label: rq2-eda-3level
survey %>%
  drop_na(assess, multitask) %>%
  count(assess, multitask) %>%
  group_by(assess) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = assess, y = prop, fill = multitask)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Assessment preparation style (3 levels)",
       y = "Proportion of students", fill = "Multitasks?") +
  theme_minimal(base_size = 13)
```

Both plots show the same pattern: the multitasking rate is visibly higher
among students who cram (~75%) than among those who don't (~52% "Gradually",
~46% "Immediately"). This motivates formally testing the association in the
next step.

## RQ2 — Step 3: Hypothesis test — Chi-squared test of independence

**H:**
$H_0$: multitasking behaviour is independent of cramming behaviour.
$H_1$: multitasking behaviour is *not* independent of cramming behaviour.

```{r}
#| label: rq2-table
cram_tab <- table(rq2_data$is_cram, rq2_data$multitask)
cram_tab
```

**A (assumptions):** The chi-squared test of independence requires (i)
independent observations (one row per student — reasonable here, no student
appears twice in the survey) and (ii) expected cell counts $\geq 5$ in every
cell of the contingency table.

```{r}
#| label: rq2-assumptions
chi_test <- chisq.test(cram_tab, correct = FALSE)
chi_test$expected
```

All four expected counts are well above 5 (minimum ≈ 51.5), so the
chi-squared approximation is valid here — there is no need to fall back on
Fisher's exact test or a Yates continuity correction.

**T & P (test statistic & p-value):**

```{r}
#| label: rq2-test
chi_test
```

```{r}
#| label: rq2-effect-size
# Effect size (Cramer's V), to accompany the p-value
n <- sum(cram_tab)
cramers_v <- sqrt(unname(chi_test$statistic) / n)
cramers_v
```

Observed: $\chi^2 = 21.35$, $df = 1$, $p \approx 3.8\times10^{-6}$,
Cramér's $V \approx 0.18$ (small-to-moderate association).

**C (conclusion with limitations):**
There is very strong evidence ($p < 0.001$) against the null hypothesis of
independence: students who cram are considerably more likely to multitask
(75.2%) than students who don't (51.8%). However:

1. This is an observational, self-selected survey sample — voluntary
   response bias means busier or more disorganised students may be
   over- or under-represented.
2. "Cram" and "multitask" are both self-reported and self-interpreted
   categories, which may not be applied consistently across respondents.
3. Association is not causation — a third factor (e.g. workload,
   procrastination tendency) could plausibly drive both behaviours.
4. Collapsing `assess` from 3 levels to 2 (Cram vs Not Cram) loses some
   information about the distinction between "Gradually" and
   "Immediately" preparers.

## RQ3

<See Assignment Briefing page>

## Final Conclusion

<See Assignment Briefing page>
